Studying public transit API query logs\\ to get an indication of travel flows

### 2016-01-07 paper.tex
Modified section 4 + figures 2 & 3

%\todo{Maybe a sub header here?} - PC: I wouldn't do a subheader...
We are particularly interested in studying the potential application of \emph{route planning} queries as an information resource for flow analysis due to the input it captures from human users.
The other features are less relevant for this purpose as they are generally polled by digital signage providers or status boards, and thus do not reflect direct human demands.
The dataset described in this paper is compiled from the \emph{Apache} \emph{access.log} files generated between October 2012 and December 2015, filtered by ``connections'' entries.
We made a copy of this dataset available in CSV format can be downloaded at \todo{\url{http://datawijs.be/apilog.tar.gz}}.
%CSV header of the final logs: datetime,requested_datetime,from_id,from_name,from_long,from_lat,to_id,to_name,to_long,to_lat,user_agent,operating_system,is_bot
Each row within this dataset contains:
\begin{enumerate}
\item A timestamp of when the request was executed,
\item The request path between two stations with longitudes and latitudes,
\item The user-agent string cfr. RFC2616,
\item The operating system of the user-agent and
\item Whether or not the user-agent is a bot.
\end{enumerate}

@@

As we were specifically interested in studying morning and evening rush hour travel that occur on weekdays, we filtered the dataset to exclude data records created outside the time range of 06:00 hrs and 10:00 hrs as well as 17:00 hrs and 21:00 hrs on weekdays, and all data records created on weekend.
Queries made by automated user-agents like search engine bots and data harvesters can be considered as valid traffic but may over represent connections between certain stations.
We have excluded such queries since we are specifically interested in studying human travel behavior.
Station names tend to contain be highly inconsistent since any text information can be parsed to the API.
This may include invalid character encoding, spelling mistakes and the use of unstandardized abbreviations.
We employ an open-source reconciliator (\url{https://github.com/irail/stations}) to reduce spelling variations and remove data records with missing information. 
Additionally, station names are automatically linked to their correspodinging geographical coordinates through this process.

\begin{figure*}
  \centering
  \includegraphics[width=14cm]{method-c.pdf}\\
  \caption{Spatial aggregation based on provincial administrative boundaries. Each station is color coded to indicate the province it will be aggregated to. Stations in major cities as well as stations outside of Belgium are aggregated to seperate distinct groups.}
  \label{fig:method-c}
\end{figure*}

\begin{figure}[h]
  \centering
  \begin{subfigure}{.2375\textwidth}
    \includegraphics[width=4cm]{method-a.pdf}\\
    \subcaption{Lines and dots\todo{Alvin}} % 4 words max
    \label{fig:method-a}
  \end{subfigure}%
  \begin{subfigure}{.2375\textwidth}
    \includegraphics[width=4cm]{method-b.pdf}\\
    \subcaption{arrows and dots\todo{Alvin}} % 4words max
    \label{fig:method-b}
  \end{subfigure}%
  \vspace{5.0pt}% To create a bit of vertical space between the 2 figure captions
  \caption{With the arrow and dots and the lines and dots you can see that TODO \todo{Alvin: can you write a caption here?}}
\end{figure}

Next, the data is aggregated so that province level flow patterns are emphasized.
\cref{fig:method} provides a diagrammatic representation of the procedure.
First, both origin and destination stations for each data record are spatially aggregated based on provincial administrative boundaries (see \cref{fig:method-a}).
Train stations in major cities are excluded from aggregation so that the volume of flow between provinces and major cities are comparable (see \cref{fig:method-b}).
If an origin or destination station is a located outside of Belgium, it is considered international travel and aggregated in a distinct group (see \cref{fig:method-c}).
The following types of flows are observable from the outcome of aggregation:

>>>>

%\todo{Maybe a sub header here?} - PC: I wouldn't do a subheader...
We are particularly interested in studying the potential application of \emph{route planning} queries as an information resource for flow analysis due to the input it captures from human users.
The other features are less relevant for this purpose as they are generally polled by digital signage providers or status boards, and thus do not reflect direct human demands.
The dataset described in this paper is compiled from the \emph{Apache} \emph{access.log} files generated between October 2012 and December 2015, filtered by ``connections'' entries.
We made a copy of this dataset available in CSV format can be downloaded at \todo{\url{http://datawijs.be/apilog.tar.gz}}.
%CSV header of the final logs: datetime,requested_datetime,from_id,from_name,from_long,from_lat,to_id,to_name,to_long,to_lat,user_agent,operating_system,is_bot
Each row within this dataset contains:
\begin{enumerate}
\item A timestamp of when the request was executed,
\item The request path between two stations and their associated geographic coordinates,
\item The user-agent string cfr. RFC2616,
\item The operating system of the user-agent and
\item A flag for whether the user-agent is a bot.
\end{enumerate}

@@

As we were specifically interested in studying morning and evening rush hour travel that occur on weekdays, we filtered the dataset to exclude data records created outside the time range of 06:00 hours and 10:00 hours as well as 17:00 hours and 21:00 hours on weekdays, and all data records created on weekend.
Queries made by automated user-agents like search engine bots and data harvesters can be considered as valid traffic but may over represent connections between certain stations.
Such queries have been excluded since we are specifically interested in studying human travel behavior.
Station names tend to contain be highly inconsistent since any text information can be parsed to the API.
This may include invalid character encoding, spelling mistakes and the use of unstandardized abbreviations.
We employ an open-source reconciliator (\url{https://github.com/irail/stations}) to reduce spelling variations and remove data records with missing information. 
Additionally, it automatically links station names to their corresponding geographical coordinates.

\begin{figure*}
  \centering
  \includegraphics[width=14cm]{method-c.pdf}\\
  \caption{Each station is color coded to indicate the province it will be aggregated to. Stations in major cities as well as stations outside of Belgium are aggregated to seperate groups.}
  \label{fig:method-b}
\end{figure*}

\begin{figure}[h]
  \centering
  \begin{subfigure}{.2375\textwidth}
    \includegraphics[width=4cm]{method-a.pdf}\\
    \subcaption{} % 4 words max
    %\label{fig:method-a}
  \end{subfigure}%
  \begin{subfigure}{.2375\textwidth}
    \includegraphics[width=4cm]{method-b.pdf}\\
    \subcaption{} % 4words max
    %\label{fig:method-b}
  \end{subfigure}%
  \vspace{5.0pt}% To create a bit of vertical space between the 2 figure captions
  \label{fig:method-a}
  \caption{Spatial aggregation based on provincial administrative boundaries diagrammatically represented. (a) Individual data records are depicted geographically with arrows connecting origins to their respective destinations. (b) The result of aggregation is a set of weighted connections between provinces.}
\end{figure}

Next, the data is aggregated so that province level flow patterns are emphasized.
\todo{Pieter: I can't seem to link the figure here} provides a diagrammatic representation of spatial aggregation on the basis of provincial administrative boundaries.
To compare the volume of flow between rural provincial areas and urbanized zones, train stations in major cities are aggregated seperately. 
If an origin or destination station is a located outside of Belgium, it is considered international travel and aggregated in a distinct group. A diagrammtic breakdown is depicted in \todo{Pieter: Same problem here...} where color and shape is used to demarcate each group.
The following types of flows are observable from the outcome of aggregation:
